\pkg{partykit}: A Modular Toolkit for Recursive Partytioning in \proglang{R}

### 2014-03-26 pkg/partykit/inst/mloss/mloss2.Rnw
ready for submission
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@559 255f4249-2263-4417-b1ea-8498779e27a7

\begin{abstract}
  The \proglang{R} package \pkg{partykit} provides a flexible toolkit for
  learning, representing, summarizing, and visualizing a wide range of tree-structured
  regression and classification models. The functionality encompasses: (a)~Basic
  infrastructure for \emph{representing} trees (inferred by any algorithm) so that
  unified \code{print}/\code{plot}/\code{predict} methods are available;
  (b)~Dedicated methods for trees with \emph{constant fits} in the leaves (or terminal nodes)
  along with suitable coercion functions to create such trees (e.g., by
  \pkg{rpart}, \pkg{RWeka}, PMML); (c)~A reimplementation of \emph{conditional inference
  trees} (\code{ctree}, originally provided in the \pkg{party} package); (d)~An extended
  reimplementation of \emph{model-based recursive partitioning} (\code{mob}, also originally
  in \pkg{party}) along with dedicated methods for trees with parametric models in the
  leaves. Here, a brief overview of the package and its design is given while more detailed
  discussions of items~(a)--(d) are available in vignettes accompanying the package.
\end{abstract}

@@

Not only trees learned within \proglang{R} can be transformed to the
proposed infrastructure but also trees from other software packages.  Either
a dedicated interface has to be created using the building blocks described
in the next section (e.g., as done for the J4.8 tree in \pkg{RWeka}) or PMML
(Predictive Model Markup Language) can be used as an intermediate exchange
format.  This is an XML standard created by an international consortium
\citep{DMG:2014} that includes a \code{<TreeModel>} tag with support for
constant-fit classification and regression trees.  The function
\fct{pmmlTreeModel} allows to read these files and represents them as
\class{simpleparty} objects in \pkg{partykit}.  The reason for not using the
\class{constparty} class as above is that the PMML format only stores point
predictions (e.g., a mean or proportion) rather than all observations of the
response that was used for tree learning.  So far, the PMML interface has
been tested with output from the \proglang{R} package \pkg{pmml} and
\proglang{SPSS}'s \pkg{AnswerTree} model.  The latter includes an
implementation of the QUEST algorithm \citep{Loh+Shih:1997}.

@@

\vspace*{-.1cm}

@@

The unified infrastructure at the core of \pkg{partykit} is especially
appealing for developers who either want to implement new tree algorithms
or represent trees learned in other systems. Here, we briefly outline
the most important classes and refer to the vignettes for more details:
\begin{description}
  \item[\class{partysplit}:] Split specification with an integer ID for the
    splitting variable, breakpoint(s), and indexes for the kids.
  \item[\class{partynode}:] Node specification with an integer ID, a
    \class{partysplit}, and a list of kids (if any) that are \class{partynode}
    objects again.
  \item[\class{party}:] Tree specification with a \class{partynode} tree
    along with a \class{data.frame} (which optionally may be empty),
    potentially plus further information about fitted values and so-called
    \class{terms} that allow to preprocess new data for predictions.
\end{description}
All classes have an additional slot for storing arbitrary information at
any level of the tree. This is exploited by \class{constparty}, \class{simpleparty},
and \class{modelparty} which store the observed response, point predictions,
and fitted parametrics models, respectively.
% Extractor functions for the various components of splits, nodes, and trees are provided.

>>>>

\begin{abstract}
  The \proglang{R} package \pkg{partykit} provides a flexible toolkit for
  learning, representing, summarizing, and visualizing a wide range of tree-structured
  regression and classification models. The functionality encompasses: (a)~basic
  infrastructure for \emph{representing} trees (inferred by any algorithm) so that
  unified \code{print}/\code{plot}/\code{predict} methods are available;
  (b)~dedicated methods for trees with \emph{constant fits} in the leaves (or terminal nodes)
  along with suitable coercion functions to create such trees (e.g., by
  \pkg{rpart}, \pkg{RWeka}, PMML); (c)~a reimplementation of \emph{conditional inference
  trees} (\code{ctree}, originally provided in the \pkg{party} package); (d)~an extended
  reimplementation of \emph{model-based recursive partitioning} (\code{mob}, also originally
  in \pkg{party}) along with dedicated methods for trees with parametric models in the
  leaves. Here, a brief overview of the package and its design is given while more detailed
  discussions of items~(a)--(d) are available in vignettes accompanying the package.
\end{abstract}

@@

Not only trees learned within \proglang{R} can be transformed to the
proposed infrastructure but also trees from other software packages.  Either
a dedicated interface has to be created using the building blocks described
in the next section (e.g., as done for the J4.8 tree in \pkg{RWeka}) or PMML
(Predictive Model Markup Language) can be used as an intermediate exchange
format.  This is an XML standard created by an international consortium
\citep{DMG:2014} that includes a \code{<TreeModel>} tag with support for
constant-fit classification and regression trees.  The function
\fct{pmmlTreeModel} allows to read these files and represents them as
\class{simpleparty} objects in \pkg{partykit}.  The reason for not using the
\class{constparty} class as above is that the PMML format only stores point
predictions (e.g., a mean or proportion) rather than all observations
from the learning sample.  So far, the PMML interface has
been tested with output from the \proglang{R} package \pkg{pmml} and
\proglang{SPSS}'s \pkg{AnswerTree} model.  The latter includes an
implementation of the QUEST algorithm \citep{Loh+Shih:1997}.

@@

@@

The unified infrastructure at the core of \pkg{partykit} is especially
appealing for developers who either want to implement new tree algorithms
or represent trees learned in other systems.

Here, we briefly outline
the most important classes and refer to the vignettes for more details:
\begin{description}
  \item[\class{partysplit}:] Split specification with an integer ID for the
    splitting variable, breakpoint(s), and indexes for the kids.
  \item[\class{partynode}:] Node specification with an integer ID, a
    \class{partysplit}, and a list of kids (if any) that are \class{partynode}
    objects again.
  \item[\class{party}:] Tree specification with a \class{partynode} tree
    along with a \class{data.frame} (which optionally may be empty),
    potentially plus further information about fitted values and so-called
    \class{terms} that allow to preprocess new data for predictions.
\end{description}
All classes have an additional slot for storing arbitrary information at
any level of the tree. This is exploited by \class{constparty}, \class{simpleparty},
and \class{modelparty} which store the observed response, point predictions,
and fitted parametrics models, respectively.
% Extractor functions for the various components of splits, nodes, and trees are provided.
